### 2023-02-12 talk/talk.tex
start to add haiti slides
haiti results

@@

\nocite{bjornstad01,grenfell04,breto19,rebeschini15,ng02,ionides21}

>>>>

\begin{frame}{Cholera in Haiti, 2010-2019}

% https://web.archive.org/web/20151106110552/http://www.ihsi.ht/pdf/projection/Estimat_PopTotal_18ans_Menag2015.pdf
%Map	Department	Capital	Area (km2)	Population	Density (Pop./km2)	Planning Region
%1	Artibonite	Gonaïves	4,987	1,727,524	350	Central
%2	Centre	        Hinche	        3,487	746,236	        210	Central
%3	Grand'Anse	Jérémie	        1,912	468,301	        240	South
%4	Nippes	        Miragoâne	1,268	342,525	        270	South
%5	Nord	        Cap-Haïtien	2,115	1,067,177	500	North
%6	Nord-Est	Fort-Liberté	1,623	393,967	        240	North
%7	Nord-Ouest	Port-de-Paix	2,103	728,807	        350	Central
%8	Ouest	        Port-au-Prince	4,983	4,029,705	810	West
%9	Sud-Est	        Jacmel	        2,034	632,601	        310	West
%10	Sud	        Les Cayes	2,654	774,976	        290	South

  % Hurricane Matthew: October 2016

\begin{columns}[T] % align columns

%\hspace{-15mm}
\begin{column}{.35\textwidth}
  \vspace{5mm}

$\bullet$  820,000 reported cases and nearly 10,000 deaths

$\bullet$  Population of 10,000,000 in 10 d\'{e}partements

    $\bullet$  Lee et al (2020) developed 3 models to guide vaccination plans:

    \vspace{1mm}

    {\bf 1}. stochastic, national \\
    {\bf 2}. ODE, spatial \\
    %d\'{e}partment \\
    {\bf 3}. stochastic, spatial
    %d\'{e}partment.

\begin{column}{.62\textwidth}
  \includegraphics[width=7.5cm]{haiti/haiti.png}

  \begin{center}
    {  \tiny \url{https://doi.org/10.5194/nhess-20-471-2020}}
  \end{center}

\end{columns}

  \vspace{4mm}

   $\bullet$ Wheeler et al (2023) continued the data analysis

\begin{frame}{Haiti cholera weekly reports 2010-2019 (log scale)}
\includegraphics[width=12cm]{haiti/Plot_Reported_Cases-1.pdf}
\end{frame}

\begin{frame}{Log-likelihoods of models and an ARMA benchmark}

  \begin{center}
\begin{tabular}{|l|rc|rc|rc|}
\hline
& Model~1& (p) & Model~2 & (p) & Model~3 & (p)
\\
\hline
Wheeler et al
& { -2731.3} & (15)
& -21957.3 & (6)
& { -17850.4} & (35)
\\
Lee et al
& -3030.9 & (20)
& -29367.4 & (6)
& -31840.8 & (29)
\\
Log-ARMA(2,1)
& -2803.7 & (4)  %% -0.5*5615.3+4
& { -18027.0} & (40)  %% -0.5*36133.9+40
& -18027.0 & (40)
\\
\hline
\end{tabular}
  \end{center}

  Model 1 (stochastic, aggregated): \\
  We add process overdispersion, and beat the ARMA benchmark.

  Model 2 (deterministic, spatial): \\
  We use a log-normal measurement model. ODE models fall far short of the ARMA benchmark.

    Model 3 (stochastic, spatial): \\
    (We fit the model using IBPF, and beat the ARMA benchmark.

@@

\nocite{bjornstad01,grenfell04,breto19,rebeschini15,ng02,ionides21,wheeler23,lee20,asfaw23arxiv}
